Mortality Notes

### 2014-11-16 proposal.Rmd
Shift to maker workflow
Added changes

---
title: "Mortality Notes"
author: "J Camac"
date: "9 July 2014"
output: pdf_document
---

@@

Mortality rates vary both spatially and temporally, and as such, are critical to accurately predict vegetation dynamics. However, few models of vegetation dynamics explicitly account for changes in mortality rates (e.g. REFS). Instead, most models used simplifed assumptions such as constant mortality (REF), no mortality (REF) or that mortality varies as a function of some arbitary set of rules (REF). These simplifcations are predominately a consequence of our limited understanding of what governs mortality rates and how it varies temporally, spatially and within and between species. This limited knowledge is partly due to ecologists placing greater emphasis on examining plant growth and how it affects community (REF) and population dynamics (REF). But also because of difficulties associated with modelling mortality. Mortality is highly variable and episodic (REF). As such, it requires individuals to be monitoried for sufficiently long periods of time, ideally from birth to death, which is problematic for long-lived species. Mortality rates are also affected by multiple biotic and abiotic factors, as well as the frequency (REF) and severity (REF) of multiple hazards (e.g. fire, wind, drought), which may or may not be dependent on one another (REF).

This complexity has led to the widespread use of phenomological models to describe patterns in plant mortality. In a review of more than 150 woody plant mortality models, Hawke (2000) found that 93% were phenomological and only 7% were mechanistic; further highlighting the paucity of process-based information about mortality. These phenological models....

These mechanistic models all assumed that mortality occured when carbon supply was insufficient for growth (e.g. Friend et al. 1997, Bossel 1986, Weinstein et al 1991, Bugmann et al 1997). As such, none incorporated differential mortality rates among species, or how different hazards may alter these rates. Phenomological model on the other hand, have attempted to explain mortality as a function of competitive interactions, abiotic conditions as well as size. One finding that appears to

How many data are needed on mortality? When and where should they be ob- tained? This is a general criticism of Friend et al. (1997), who claim these problems render valida- tion of mechanistic models useless. Also, data sources for mortality testing are often flawed. Bugmann (1996) points out that monitoring pro- grammes are usually shorter than many plants’ lifespans, while palaeo data does not provide the necessary resolution. And the potential variation between sizes, classes, ages etc. is so high that even a large sample will not provide the statistical significance required.

@@

>>>>

---
title: "Mortality Notes"
author: "J Camac"
date: "9 July 2014"
output: pdf_document
bibliography: data/refs.bib
csl: downloads/style.csl
---

@@

[@kleinbaum_survival_2005]

Mortality rates vary both spatially and temporally, and as such, are critical to accurately predict vegetation dynamics. However, few models of vegetation dynamics explicitly account for changes in mortality rates. Instead, most models used simplifed assumptions such as constant mortality rates (Thorton 2007m White 2000, Delbart et al 2010) or that mortality varies as a function of some arbitary set of rules or thresholds (Sitch et al 2003, Sato et al 2007). In many cases, mortality is not simulated at all (Hawkes 2000). These simplifications are predominately a consequence of our limited understanding of what governs mortality rates and how it varies temporally, spatially and within and between species (Hawkes 2000, McDowell et al 2011). This limited knowledge is partly due to ecologists placing greater emphasis on examining plant growth and how it affects community and population dynamics (Hawke 2000). But also because of difficulties associated with modelling mortality (Hawke 2000). Mortality is highly variable and episodic (Burgmann 1994). As such, it requires individuals to be monitoried for sufficiently long periods of time, ideally from birth to death, which is problematic for long-lived species (maybe Bugman 1996). Mortality rates are also affected by multiple biotic and abiotic factors, as well as the frequency and severity of multiple hazards (e.g. fire, wind, drought), which may or may not be dependent on one another (McDowell 2011).

This complexity has led to the widespread use of phenomological models that attempt to describe patterns in plant mortality as a function of correlations with various covariates. In a review of more than 150 woody plant mortality models, Hawke (2000) found that 93% were phenomological and only 7% were mechanistic. Such phemenological models have been used to document how mortality varies as a function of competition (REF), climate (REF), age (REF) and size (REF). The effect of size has been one

These mechanistic models all assumed that mortality occured when carbon supply was insufficient for growth (e.g. Friend et al. 1997, Bossel 1986, Weinstein et al 1991, Bugmann et al 1997; Fisher et al 2010). As such, none incorporated differential mortality rates among species, or how different hazards may alter these rates. Phenomological model on the other hand, have attempted to explain mortality as a function of competitive interactions, abiotic conditions as well as size. One finding that appears to

And the potential variation between sizes, classes, ages etc. is so high that even a large sample will not provide the statistical significance required.

@@

# REFERENCES
